Sampling Strategies in Decisions from Experience

### 2022-01-23 manuscript.Rmd
Update Section 2

Since it was stated that the inputs to the decision making process can be organized in a set of choice alternatives, we can adapt (1) for the case of a choice between $1, ..., n$ prospects to obtain:

$$
\iota : I' \mapsto I, \ \text{where} \ I = \{(\Omega_1,  \mathcal{F}_1, P_1), ..., (\Omega_n,  \mathcal{F}_n, P_n)\}
\; .
\tag{3}
$$

@@

Note that a selection criterion may refer to a property for which the raw input, i.e., the symbols, their organization, and the information they encode, do not yet provide an explicit description.
Even more so, it has been repeatedly stated that for quite many decisions the mind faces, it is simply not possible to draw on direct descriptions of the properties, since they are part of the distal environment  [e.g., @fiedlerExplainingSimulatingJudgment1996; @hertwigDescriptionexperienceGapRisky2009; see also @brunswikRepresentativeDesignProbabilistic1955; @brunswikPerceptionRepresentativeDesign1956; @kozyrevaInterpretationUncertaintyEcological2021].
In either case, some form of description must be inferred from whatever information the input provides about the property in question.
Consider first the counterfactual case of the prospect whose tuple is not part of the distal environment but is directly described. 
Since all information about the prospect is given, one may in principle consider the input complete, yet, the tuple implies properties of the prospect which are not explicitly described.
For example, to obtain the description of the prospect's weighted average, still a calculus according to the algebraic rules described in the preceding paragraph must be carried out. 
In general, in cases like the one just considered, where the properties are not distal and all information about them is available, property inferences take the form of a mere transformation, where the symbols from the input are ordered and/or combined according to some transformation rules [cf. @marrVisionComputationalInvestigation1982, p. 20].

In natural world decisions, however, the properties of choice alternatives are often distal and amenable only through their probabilistic relation to proximal cues or samples.
Such a proximal information basis, obtained by experience or, more technically, by *sampling*, is neither consistent nor complete but requires the mind to draw inductive inferences to arrive at property descriptions which then, by definition, are only probable. 
In turn, the "achievement" [@brunswikOrganismicAchievementEnvironmental1943, p. 255] of the mind, i.e., highly probable descriptions, depends in part ... [but see @fiedlerInformationSamplingAdaptive2006, for a selection of papers which goes beyond the mere discussion of statistically representative samples].

To adapt (3) such that it fits to the sampling, as it is done for decisions from experience [@hertwigDecisionsExperienceEffect2004], we define sampling as follows:

@@

(1) $X$ is a measurable function $$X: (\Omega, \mathcal{F})  \mapsto (\Omega', \mathcal{F'}) \; ,$$ where $\Omega'$ is a set of real numbered values $X$ can take and $\mathcal{F'}$ is a set of subsets of $\Omega'$. I.e., $\Omega$ maps into $\Omega'$ such that correspondingly each subset $A' \in \mathcal{F'}$ has a pre-image $X^{-1}A' \in \mathcal{F}$, which is the set $\{\omega \in \Omega: X(\omega) \in A'\}$ [@kolmogorovFoundationsTheoryProbability1950, p. 21].

>>>>

Since it was stated that the inputs to the decision making process can be organized in a set of choice alternatives, we can adapt (1) for the case of a choice between $1, ..., k$ prospects to obtain:

$$
\iota : I' \mapsto I, \ \text{where} \ I = \{(\Omega_1,  \mathcal{F}_1, P_1), ..., (\Omega_k,  \mathcal{F}_k, P_k)\}
\; .
\tag{3}
$$

@@

Note that a selection criterion may refer to a property for which the raw input, i.e., the symbols, their organization, and the information they encode, do not yet provide an explicit description.
Even more so, it has been repeatedly stated that for quite many decisions the mind faces, it is simply not possible to draw on direct descriptions of the properties, since they are part of the distal environment  [e.g., @brunswikOrganismicAchievementEnvironmental1943; @fiedlerExplainingSimulatingJudgment1996; @hertwigDescriptionexperienceGapRisky2009; see also @brunswikRepresentativeDesignProbabilistic1955; @brunswikPerceptionRepresentativeDesign1956; @kozyrevaInterpretationUncertaintyEcological2021].
In either case, some form of description must be inferred from whatever information the input provides about the property in question.
Consider first the counterfactual case of the prospect whose tuple is not part of the distal environment but is directly described. 
Since all information about the prospect is given, one may in principle consider the input complete, yet, the tuple implies properties of the prospect which are not explicitly described.
For example, to obtain the description of the prospect's weighted average, still a calculus according to the algebraic rules described in the preceding paragraph must be carried out. 
In general, in cases like the one just considered, where the properties are not distal and all information about them is available, property inferences take the form of a mere transformation, where the symbols from the input are ordered and combined according to some transformation rules [cf. @marrVisionComputationalInvestigation1982, p. 20].

In natural world decisions, however, the properties of choice alternatives are often distal and amenable only through their probabilistic relation to proximal cues or samples, which provide the input to the decision making process [cf. @brunswikOrganismicAchievementEnvironmental1943].
Such a proximal information basis is obtained by experience or, more technically, by *sampling*, and is neither consistent nor complete but varies in its degree of representativeness. 
In turn, the mind is not merely required to transform the inputs but to draw inductive inferences to arrive at property descriptions which then, by definition, are only probable. 
Importantly, the probability, or accuracy, of the description depends on both the sampling process that determines the representativeness of the sample and the use of the probabilistic information that is entailed in the sample [e.g., @fiedlerBewareSamplesCognitiveecological2000].

To adapt (3) for the case of decisions from experience [@hertwigDecisionsExperienceEffect2004], the tuples $(\Omega, \mathcal{F}, P)$ must be replaced by samples, which in this paper are treated as collections of independent and identically distributed (i.i.d.) random variables $X$.^[Note that this paper considers the counterfactual case where samples are random, although it has been stated that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. To account for non-random sampling, e.g., in the form of sampling biases [e.g., @fiedlerBewareSamplesCognitiveecological2000] or dynamic information structures [e.g., @cohenEffectPerceivedPatterns2021; @plonskyRelianceSmallSamples2015], the i.i.d-assumption would need to be dropped, which is out of the scope of this paper.]
Specifically, random variables are defined as a measurable function

$$
X: (\Omega, \mathcal{F})  \mapsto (\Omega', \mathcal{F'}) 
\; ,
\tag{4}
$$

where $\Omega'$ is a set of real numbered values $X$ can take and $\mathcal{F'}$ is a set of subsets of $\Omega'$. 
I.e., $\Omega$ maps into $\Omega'$ such that correspondingly each subset $A' \in \mathcal{F'}$ has a pre-image $X^{-1}A' \in \mathcal{F}$, which is the set $\{\omega \in \Omega: X(\omega) \in A'\}$ [@kolmogorovFoundationsTheoryProbability1950, p. 21].

$$
\begin{aligned}
  & \iota : I' \mapsto I, \ \text{where} \ I = \{X_{1i}, ..., X_{ki}\} \\
  & i \in \{1, ..., N\} 
  \; .
\end{aligned}
\tag{5}
$$

@@
